environmental_data_preperation

### 2018-10-25 src/biological_data_prep.Rmd
bottom depth extraction

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
Th
```{r pre-install packages, message=FALSE}
packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp", "raster", "tibble", "useful")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

>>>>

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
Th
```{r pre-install packages, message=FALSE}
packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp", "raster", "tibble", "useful", "marmap")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

### 2018-10-25 src/environmental_data_preperation.Rmd
feck all progress :(
bottom depth extraction

>>>>

# Bathymetric

bottom depth from [GEBCO](www.gebco.net)

```{r}
library(ncdf4)
gebco <- nc_open("../data/env/GEBCO_2014_2D_-70.0_34.0_-43.0_70.0.nc")
gebco
library(marmap)
bat <-readGEBCO.bathy("../data/env/GEBCO_2014_2D_-70.0_34.0_-43.0_70.0.nc")
summary(bat)
```

create pretty custom color palette and then plot the bathy laye
```{r}
blues <- colorRampPalette(c("lightblue", "cadetblue2", "cadetblue1", "white"))
plot(bat, n = 1, image = TRUE, bpal = blues(100))
```
lovely jubbely

add the species data...
```{r}
data_aea <-read.csv( "../output/bio/data_aea_cell_amo_nao_bins_env.csv", header = TRUE) #just for canada pc as not loaded
plot(bat, n = 1, image = TRUE, bpal = blues(100))
points(data_aea$decimalLongitude, data_aea$decimalLatitude, pch = 20, col = "yellow") #note these are both the wgs84 images
```
and now get depth... note all the points and bathy data is wgs!

```{r}
data_aea$bottom_depth <- get.depth(bat, data_aea[ ,5:4], locator = FALSE)
data_aea$bottom_depth <- data_aea$bottom_depth[, 3] #this step is because get.bath actually produces it's own dataframe with long, lat, and depth. This is extracting the depth column and deplacing data_aea$bottom_depth with the correct data.
head(data_aea)
```

might be easier just to work form a TIF though...

First project the tiff to equal area
```{r}
library(raster)
geb <- "../data/env/gebco.tif"
mask1 <- "../data/env/aea.tif" #just for canada pc as file isn't loaded
aea <- raster(mask1)
gebr <- brick(geb)
gebr
aea
geb
geb_aea <- projectRaster(geb, aea) 
```

```{r GEBCO raster extraction}
library(rgdal)
library(sp)
gebco_coordinates(data_aea) <- c("decimalLongitude", "decimalLatitude") 
nafo_zones <- readOGR(dsn=path.expand("../data/bio/nafo_zones"), layer = "nafo_zones_wgs84") #this loads the shapefile
proj4string(adddata) <- proj4string(nafo_zones) #tells R that the occurrence data is the same projection as the shapefile
adddata$nafo_zone <- over(adddata, nafo_zones)$ZONE #ZONE is where the zone data is held in teh shapefiles' attributes
head(adddata) #just to check it worked
```
